---
title: "Spanish Email Network"
author: "Yijia Lin, Bradley McKenzie and Diego Paroli"
date: "`r Sys.Date()`"
output: html_document
---

# Libraries

```{r, message=FALSE, warning=FALSE}
rm(list = ls())
library(tidyverse)
library(httr2)
library(igraph)
library(visNetwork)
library(tidygraph)
library(ggraph)
library(here)
```

# Get the data

This network is a directed and unweighted social communication network, which represents the exchange of emails among members of the Rovira i Virgili University in Spain, in 2003.

Source: [Netzschleuder](https://networks.skewed.de/net/uni_email).

### Nodes and links

```{r, message=FALSE, warning=FALSE}
nodes <- read_csv(here("network_data", "nodes.csv"))
links <- read_csv(here("network_data", "edges.csv"))
```

```{r}
head(nodes)
head(links)
```

Within the `nodes` dataframe, the column `# index` is the one indicating the index of the nodes,
the column `name` is the one indicating their corresponding name (although already anonymised), and the `_pos` column represents the coordinates of each node in a 2D space, typically used for visualization or layout in graph-related tasks. In the `links` dataframe, the column `# source` and `target` indicate the 2
nodes forming a link

### Graph:

```{r}
graph <- graph_from_data_frame(links, directed = TRUE, vertices = nodes)
graph
```
